R Notebook

### 2025-03-10 Scripts/authoritarian_only_latex_2.Rmd
Major changes to binary models (includes incumbent losses now)

There are two problems, however. First, the empirical record on the relationship between manipulation and protest been mixed, illustrating the limitations of current theory. [FILL IN CITES HERE] Second, the expected relationship between manipulation and protest under the regime-weakness and regime-strength models is complicated by the emergence of a grievance-based model of electoral protest generally. This view, which has considerable empirical support, finds that electoral protests are largely driven by citizen grievances\textemdash not the narrowly strategic calculations about the likelihood of overthrowing the government which are assumed in the regime-weakness model. Under the grievance framework, when citizens have many individual grievances against the incumbent, a highly manipulated election can serve as a coordination point for collective action [@tucker2007].

@@

[Somewhere cite this article: https://doi.org/10.1177/0032321716674024 on how 'toleration' is really just slow attrition of protest movements.]

@@

> H1: Increasing election fraud makes protest more likely, with the size of this effect decreasing with incumbent vote-share.

@@

Simpser's foundational argument in the regime-strength model holds that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p. 80]\textemdash that is, of its political strength. Additional research focusing on the principal-agent dyanmics inherent in election manipulation reinforces the idea that manipulation carries such costly signals, such that more extensive electoral manipulation indicates greater regime strength. This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].

@@

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1 & inc.lose == 0)

@@

While it is possible that the theoretical mechanisms considered here may be at work in democracies as well, democracies are excluded from this analysis for two reasons. First, the theory presented above is derived from prior work on authoritarian elections; the factors that drive the theory are expected to be weaker in democratic regimes. As regimes trend toward liberal democracy, patronage consolidation should decline and rule of law should increase by definition. Consequently, the theory expects that fraud should be rare in such cases, even in the absence of a protest deterrent: incumbents simply lack the tools. At the same time, the more liberal-democratic a state becomes, the more alternatives their are to protest as a means to check the executive. More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest\textemdash the electoral authoritarian regimes\textemdash where we should expect to see a more pronounced role for protest risk in curtailing election fraud under the deterrence logic. The case selection strategy here is thus one that both fits the theory and makes for a strong test of the classic fraud-protest model.

In all models, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the population of interest: protest following an election in which the incumbent claimed victory, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election is necessary to avoid conflating pro- and anti-incumbent protest. However, this design does influence the interpretation of the results from these models. Threats to inference from this section are evaluated in more detail in the discussion section.] The full dataset includes 445 election-year observations across 81 non-democratic countries. The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.^[Some additional tests in the appendix include cases where incumbents accepted defeat, bringing the total number of observations to 659.]

For models of protest-wave duration, only observations where post-election protests were initiated can be included. For these models, therefore, I limit that analysis to only those cases with a positive number of post-election protests. Of the 445 elections in which the incumbent claimed victory, only 156 fall into this more limited dataset. Brief further analyses of the seven cases in which incumbents were ousted, as well as cases where incumbents stepped down after the election, are also included in the discussion and appendix.

@@

The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0 & inc.lose == "0") %>% nrow()` of the 445 elections, no protests are recorded. Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration. As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration. This distribution\textemdash with large numbers of zeroes\textemdash is in line with the theoretical proposition that most fraudulent elections do not produce signals of weakness that prompt opposition protest. Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression. Protest wave duration is modeled using Cox proportional hazards models.

@@

p.hist.events <- ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()

p.hist.duration <- ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()

p.hist.fraud <- ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()

 ggplot(nelda2.sub %>% filter(inc.lose == "0"), aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()

@@

The models incorporate several potential confounding variables, in addition to the *election fraud* variable described above. The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance. The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.}

As identified in prior research, other factors that may influence both fraud and protest are included as controls variables. The repressive capacity of the government [@Davenport2019] is measured using the physical integrity index from V-Dem (v2x_clphy), and its square term. Indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), de facto *judicial independence* [@Linzer2015] and the available of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem) are also employed. Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019]. All models include *legislative constraints* on the executive, a V-Dem variable which captures the extent to which the legislative body in the country acts independently in practice to limit executive power. Additional V-Dem variables capture *privatization* of the economy ('v2clstown') and *bureaucratic professionalization* ('v2stcritrecadm'), both of which might give incumbents influence over election fraud and subsequent protest. Lastly, a V-Dem measure the *autonomy of election managment bodies* ('v2elembaut') is included.

Several binary control variables are also included. First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest. Another NELDA variable indicates if the incumbent executive is not running due to *term limits* [@hale_patronal_2014]. I include dummy variables that provide a categorical evaluation of state capacity. These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap). If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*. If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*. States that rely primarily on taxes on property or economic transactions are coded as *tax states*. This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018]. Since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.

@@

```{r include=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

model.margin.binary <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   + v2elirreg.inv +  +    winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary)

model.margin.binary.int <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary.int)

model.margin.binary_noc <- glm(any_protest ~ lji.lag + legcon.1lag +  altinf.1lag  +  + osorg.1lag   + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary_noc)

model.binary_noc <- glm(any_protest ~  v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential , family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.binary_noc)

# Plots

####Margin

#plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
#plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()

plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.32, 2.2]"), show.data = T, colors = "gs")
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()

@@

```{r logit-table, echo=FALSE}
models2 <- list()
models2[['Model 1']] <- model.margin.binary
models2[['Model 2']] <- model.binary_noc
models2[['Model 3']] <- model.margin.binary.int
#models2[['Model 6']] <- model.margin.binary_noc

modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', gof_omit = c("Log|RMSE"), coef_map = c("(Intercept)" = "(Intercept)",
  "lji.lag" = "Judical independence (lag)",                                                                 "legcon.1lag" = "Leg. constraints (lag)",                                                                 "gdpgro.1lag" = "GDP growth rate (lag)",                                                                   "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)",
                 "inc.termlimit.num" = "Incumbent term-limited", 
                 "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", 
                 "rentierstate1" = "Rentier state", 
                 "taxstate1" = "Tax state",

                 "unscheduled_election1" = "Off-schedule election",
                 "n.events.pre" = "Number of pre-election protests", 
                 "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", 
                 "regional_success_lag" = "Regional protest diffusion",
                                  "presidential1" = "Presidential election",

                 "median.participants" = "Median protest participants",
                  "winner.share.upd" = "Incumb. vote-share",
                  "v2elirreg.inv" = "Election fraud",
                "v2elirreg.inv:winner.share.upd"= "Election fraud:Incumb. vote-share")
                 )

Table \@ref(tab:logit-table) shows the results for Hypothesis 1. Model 1 is a base model, where incumbent vote-share and *intentional irregularities* are standalone variables. In keeping with the ambiguous findings in the prior literature, there is no significant relationship between fraud severity and the likelihood of protest initiation. The incumbent's claimed margin of victory is negatively associated with protest initiation, though this relationship falls short of conventional statistical significance. Several control variables behave as expected; for instance, protests are more likely to occur after presidential elections and when there have been prior electoral protests in the region. Model 2 includes the interaction between *intentional irregularities* and incumbent vote-share, with no controls other than a dummy variable for *presidential elections*. This allows for the largest possible sample size, and the results are consistent with those of the full model with controls.

>>>>

There are two problems, however. First, the empirical record on the relationship between manipulation and protest been mixed, illustrating the limitations of current theory [@Hyde2014;@daxecker2012;@Hafner-Burton2016;@Lankina2017;@ananyevDictatorsSignalStrength2022]. Second, the expected relationship between manipulation and protest under the regime-weakness and regime-strength models is complicated by the emergence of a grievance-based model of electoral protest generally. This view, which has considerable empirical support, finds that electoral protests are largely driven by citizen grievances\textemdash not the narrowly strategic calculations about the likelihood of overthrowing the government which are assumed in the regime-weakness model. Under the grievance framework, when citizens have many individual grievances against the incumbent, a highly manipulated election can serve as a coordination point for collective action [@tucker2007].

@@

@@

> H1: Increasing election fraud makes protest more likely, with the size of this effect decreasing with incumbent-party vote-share.

@@

Simpser's foundational argument in the regime-strength model holds that manipulated elections can send costly signals of the incumbent's "capacities, resources, [and] dispositions" that will help it retain power in the future [@simpser2013, p. 80]\textemdash that is, of its political strength. Additional research focusing on the principal-agent dynamics inherent in election manipulation reinforces the idea that manipulation carries such costly signals, such that more extensive electoral manipulation indicates greater regime strength. This research trend emphasizes the degree to which political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].

@@

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)

@@

While it is possible that the theoretical mechanisms considered here may be at work in democracies as well, it is likely that these mechanisms are weaken as democratization advances. As regimes trend toward liberal democracy, patronage consolidation should decline and rule of law should increase by definition. These changes simultaneously make it more difficult for incumbents to manipulate elections and easier for opponents to engage in electoral protest. At the same time, the more liberal-democratic a state becomes, the more alternatives there are to protest as a means to check the executive. More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest\textemdash the electoral authoritarian regimes\textemdash where we should expect the deterrence logic to operate under a regime-weakness model. Finally, as a matter of measurement, protest duration is likely to be an increasingly imperfect measure of incumbent strength in more liberal-democratic settings, where protest is more commonplace and less immediately threatening to the regime. The case selection strategy here, excluding electoral and liberal democracies, is thus one that both fits the theory and makes for a stronger test of the regime-weakness model. Nevertheless, an expanded sample that includes electoral democracies provides similar results in the appendix. In total, there are 659 elections included in the dataset, across 100 countries. The date range is limited by the availability of data on protest from ECAV, which ranges from 1989 to 2012.

For models of protest-wave duration, as discussed in more detail below, I include only observations where a positive number of protests occurred. Only 156 fall into this more limited dataset. Brief further analysis of the seven cases in which incumbents were successfully ousted is also included in the discussion and appendix.

@@

The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nelda2.sub %>% nrow()` elections, no protests are recorded. Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration. As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration. These distributions\textemdash extensive election fraud combined with large numbers of non-protest\textemdash provides some early indication that the regime-weakness model has limited support. Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression. Protest wave duration is modeled using Cox proportional hazards models.

@@

p.hist.events <- ggplot(nelda2.sub, aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()

p.hist.duration <- ggplot(nelda2.sub, aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()

p.hist.fraud <- ggplot(nelda2.sub, aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()

 ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()

@@

It is necessary to condition the models on whether or not the incumbent party claimed victory in the election, due to the structure of the vote-share data in V-Dem. Vote-share data is provided for the winning party without reference to its position as incumbent or challenger. From the point of view of those considering anti-incumbent protest, for instance, the information conveyed by a large opposition vote-share is much different from that conveyed by a large winning vote-share. Consequently, in all models, I condition for whether the incumbent party lost, based on coding in NELDA. In the binary models of Hypothesis 1, I control for incumbent loss as a standalone variable or in interaction with the main explanatory variables.

For the models of protest duration, I exclude all cases where the incumbent lost the election outright. This limits the observations to those that fit the population of interest: protest following an election in which the incumbent claimed victory, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election avoids introducing measurement error in the dependent variable, due to a quirk of the ECAV dataset, which codes the 'side' of the protest (pro- or anti-government) as missing during the period between an election and the swearing in of a new government in cases where the incumbent loses. Including these cases thus runs the risk of conflating pro- and anti-incumbent protest, along with cases where protesters goals are genuinely unknown. However, excluding them also influences the interpretation of the results from these models. Threats to inference from this section are evaluated in more detail in the discussion section.]

The models incorporate several potential confounding variables, in addition to the *election fraud* variable described above. The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance. The winning party's *vote-share* is also taken from the V-Dem dataset.^[This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia, where links to official sources were available, and Psephos.]

As identified in prior research, other factors that may influence both fraud and protest are included as controls variables. The repressive capacity of the government [@Davenport2019] is measured using the physical integrity index from V-Dem (v2x_clphy), and its square term. Indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), de facto *judicial independence* [@Linzer2015] and the available of *alternative sources of information* ('v2x_freexp_altinf' from V-Dem) are also employed. Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019]. All models include *legislative constraints* on the executive, a V-Dem variable which captures the extent to which the legislative body in the country acts independently in practice to limit executive power.

In addition, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby [@Bunce2010;@Beissinger_2007]. Such a dynamic could plausibly increase the likelihood of protest in a nearby country, while also reducing the incumbent's use of fraud. I control for this using each country's geopolitical region as coded in Teorell et al [-@qog_teorell], in conjunction with binary variables indicating a successful protest movement from NELDA. A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down. The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'. All told there are 14 successful electoral protests in the dataset: Bulgaria 1990, Burkina Faso 1997, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Mali 2002, Georgia 2003, Ukraine 2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]

Several other binary control variables are also included. First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest. Another NELDA variable indicates if the incumbent executive is not running due to *term limits* [@hale_patronal_2014]. I include dummy variables that provide a categorical evaluation of state capacity. These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap). If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*. If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*. States that rely primarily on taxes on property or economic transactions are coded as *tax states*. This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018]. Since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

model.binary_noint <- glm(any_protest ~ inc.lose + v2elirreg.inv + winner.share.upd, data = nelda2.sub, family = binomial(link = "logit"))

model.binary.base <-  glm(any_protest ~ winner.share.upd*v2elirreg.inv*inc.lose + presidential, data = nelda2.sub, family = binomial(link = "logit"))

model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre + rentierstate + taxstate + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)

plot_binary_margin_typ_n <- sjPlot::plot_model(model.binary.full, type = "pred", terms = c("v2elirreg.inv", "winner.share.upd [.35, .75]", "inc.lose"), show.data = T)
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Electoral manipulation", y = "Predicted probability", title = "Predicted probability of protest", color = "Winner's vote-share") + theme_bw()
```

```{r logit-table, echo=FALSE}
models2 <- list()
models2[['Model 1']] <- model.binary_noint
models2[['Model 2']] <- model.binary.base
models2[['Model 3']] <- model.binary.full
#models2[['Model 6']] <- model.margin.binary_noc

modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', gof_omit = c("Log|RMSE"), coef_map = c("(Intercept)" = "(Intercept)",
             "inc.lose1" = "Incumbent loss",
             "v2elirreg.inv" = "Election fraud",
             "winner.share.upd" = "Winner vote-share",
"winner.share.upd:v2elirreg.inv" = "Winner vote-share:Election fraud",
"winner.share.upd:inc.lose1" = "Winner vote-share:Inc. loss",
"v2elirreg.inv:inc.lose1" = "Election fraud:Inc. loss",
"winner.share.upd:v2elirreg.inv:inc.lose1" = "Election fraud:Winner vote-share:Inc. loss",
  "lji.lag" = "Judical independence (lag)",                                                                 "legcon.1lag" = "Leg. constraints (lag)",                                                                 "gdpgro.1lag" = "GDP growth rate (lag)",                                                                   "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)",
                 "inc.termlimit.num" = "Incumbent term-limited", 
                 "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", 
                 "rentierstate1" = "Rentier state", 
                 "taxstate1" = "Tax state", 
                "presidential1" = "Presidential",
                 "unscheduled_election1" = "Off-schedule election",
                 "n.events.pre" = "Number of pre-election protests", 
                 "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", 
                 "regional_success_lag" = "Regional protest diffusion",

                 "median.participants" = "Median protest participants",

                "v2elirreg.inv:winner.share.upd"= "Election fraud:Winner vote-share")
                 )

@@

Table \@ref(tab:logit-table) shows the results for Hypothesis 1. Model 1 is a base model, where incumbent loss, winner vote-share, and *election fraud* are standalone variables. In keeping the general (though inconsistent) findings in the prior literature, there is a significant positive relationship between fraud severity and the likelihood of protest initiation. The winner's claimed margin of victory is negatively associated with protest initiation. Model 2 includes the three-way interaction between *election fraud*, incumbent loss, and winning vote-share, with no controls other than a dummy variable for *presidential elections*. This allows for the largest possible sample size, and the results are consistent with those of the full model with controls.
